---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.  -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gregRy

<!-- badges: start -->
<!-- badges: end -->

The goal of `gregRy` is to make the GREGORY estimator easily available to use. 

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(pdxTrees)
library(ggplot2)
```

## Installation

`gregRy` can be installed from CRAN with:

```{r, eval = FALSE}
install.packages("gregRy")
```


The development version of `gregRy` is available from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("olekwojcik/gregRy")
```

## Example Computations

The package `gregRy` does not contain a dataset, which is why our example utilizes the package [pdxTrees](https://github.com/mcconvil/pdxTrees)

### GREGORY

```{r, warning=FALSE, message=FALSE}
library(gregRy)
set.seed(13)
# load and wrangle data

# Overall dataset to create estimates with
# Includes response variable and predictors


family_filter_data <- get_pdxTrees_parks() %>%
  as.data.frame() %>%
  drop_na(DBH, Crown_Width_NS, Tree_Height) %>%
  filter(Condition != "Dead") %>%
  select(UserID, Tree_Height, Crown_Width_NS, DBH, Condition, Family) %>%
  group_by(Family) %>%
  summarize(count = n())

dat <- get_pdxTrees_parks() %>%
  as.data.frame() %>%
  drop_na(DBH, Crown_Width_NS, Tree_Height) %>%
  filter(Condition != "Dead") %>%
  select(UserID, Tree_Height, Crown_Width_NS, DBH, Condition, Family) %>% 
  left_join(family_filter_data, by = "Family") %>%
  filter(count > 4) %>%
  select(UserID, Tree_Height, Crown_Width_NS, DBH, Condition, Family)

dat_s <- get_pdxTrees_parks() %>%
  as.data.frame() %>%
  drop_na(DBH, Crown_Width_NS, Tree_Height) %>%
  filter(Condition != "Dead") %>%
  select(UserID, Tree_Height, Crown_Width_NS, DBH, Condition, Family) %>%
  left_join(family_filter_data, by = "Family") %>%
  filter(count > 4) %>%
  select(UserID, Tree_Height, Crown_Width_NS, DBH, Condition, Family) %>%
  group_by(Family) %>%
  slice_sample(prop = 0.25) %>%
  ungroup()

dat_est <- dat %>%
  filter(Family == "Pinaceae")
predictors <- c("Crown_Width_NS", "DBH")

dat_x_bar <- dat %>%
  group_by(Family) %>%
  summarize(across(predictors,
                   mean)) %>%
  pivot_longer(!Family,
               names_to = "variable",
               values_to = "mean")

dat_count_est <- dat %>%
  group_by(Family) %>%
  summarize(count = n())

# Create dataset of proportions using estimation and resolution

dat_prop <- left_join(dat, dat_count_est, by = "Family") %>%
  group_by(Condition, Family) %>%
  summarize(prop = n()/count) %>%
  distinct() %>%
  ungroup()

# Create dataset of means of 'pixel' data

dat_x_means <- get_pdxTrees_parks() %>%
  as.data.frame() %>%
  drop_na(DBH, Crown_Width_NS, Tree_Height) %>%
  dplyr::summarize(DBH = mean(DBH), Crown_Width_NS = mean(Crown_Width_NS),
            Tree_Height = mean(Tree_Height))

dat_x_bar_new <- dat_x_bar %>%
  filter(variable == "Crown_Width_NS") %>%
  mutate(Crown_Width_NS = mean) %>%
  select(Family, Crown_Width_NS)
```

To use GREGORY, we need 3 different datasets. 

The first dataset is the overall data:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
print(head(dat))
```

The second dataset is the means of the predictors at the estimation level (Family estimates):

```{r, warning=FALSE, message=FALSE, echo=FALSE}
print(head(as.data.frame(dat_x_bar_new)))
```

The third dataset is contains both the resolution and estimation, with the proportion of resolution in the given estimation unit:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
print(head(as.data.frame(dat_prop)))
```

```{r, warning=FALSE, message=FALSE}
# Create GREGORY estimates
x1 <- gregory_all(plot_df = dat_s %>% drop_na(),
            resolution = "Condition",
            estimation = "Family",
            pixel_estimation_means = dat_x_bar_new,
            proportions = dat_prop,
            formula = Tree_Height ~ Crown_Width_NS,
            prop = "prop")
print(x1)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
ggplot(data = x1, aes(y = estimate, x=Family)) +
  geom_col(fill="steel blue") + 
  labs(title = "GREGORY Estimates of Tree Height using Crown Width", x = "Tree Family", y="Estimate") + 
  coord_flip() + 
  theme_minimal() +
  theme(axis.text.y = element_text(size=8))
```

### GREG

```{r, warning=FALSE, message=FALSE}
# Create GREG estimates
x2 <- greg_all(plot_df = dat_s %>% drop_na(),
         estimation = "Family",
         pixel_estimation_means = dat_x_bar_new,
         formula = Tree_Height ~ Crown_Width_NS)
print(x2)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
ggplot(x2, aes(y = estimate, x=Family)) + 
  geom_col(fill="steel blue") +
  labs(title = "GREG Estimates of Tree Height using Crown Width",  x = "Tree Family", y="Estimate") +
  theme_minimal() +
  theme(axis.text.y = element_text(size=8)) +
  coord_flip()
```
